---
title: "coding with base R"
author: "Eli Schulman"
date: "11/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Coding with base R

Vectors

```{r}
# create a numeric vector with default values (0)
my_vector <- vector(mode = "numeric", length = 100)

# default for logical is "False"
my_second_vector <- vector(mode = "logical", length = 10)

my_vector
my_second_vector
```

```{r}
# show length, class

length(my_vector)
class(my_second_vector)
```

identifying numbers in a vector

```{r}
vector_1 <- c(1:50)

vector_1[5]
vector_1[10:20]
vector_1[-3]
```

Create a Matrix
```{r}
matrix_1 <- matrix(data = 1:100, nrow = 5, ncol = 20)

matrix_1
```

```{r}
# use dim() to return number or rows and number of columns
dim(matrix_1)

class(matrix_1)
```

Indexing

```{r}
#specify just a row or just a column to retried that whole list 
matrix_1[1, ]

#specify row and column to return value
matrix_1[2,3]

# Index and replace
matrix_1[1, 5] <- 1000000

matrix_1
```

For Loops
```{r}
for (i in 1:10){
  print(i)
}

```

Generating Random Numbers
```{r}
# use rnorm to pull random numbers from a theoretical normal dist.
rnorm(n = 5, mean = 10, sd = 2)
```

Combing everything to do the following:
- use 'numeric' to create empty vector
- use 'for' to iterate through 
- use '[ ]' to access each spot and do something 
- use 'rnorm' to generate a random number for that spot

$$N_{t}=N_{t-1} + r\left(1 - \frac{N_{t-1}}{K}\right)$$
$$
r \sin norm(\mu = 1.5, \sigma = 0.8)
$$

Creating the model with specified parameters
Model 1
```{r}
# setup model parameters

K <- 100
nsteps <- 100
mu <- 1.5
sigma <- 0.8

# setup model objects

N = vector(mode = "numeric", length = nsteps)
# assign starting value of 25 (would be 0 otherwise)
N[1] = 25

#Model it starting at point 2 since we already gave a value for point 1
for (t in 2:nsteps){
  r <- rnorm(n = 1, mean = mu, sd = sigma)
  N[t] <-  N[t-1] + r + (1 - N[t - 1] / K)
}

# visualize with a plot

plot(N)
```
Model 2
```{r}
# setup model parameters

K <- 100
nsteps <- 100
mu <- .4
sigma <- 1.5

# setup model objects

N = vector(mode = "numeric", length = nsteps)
# assign starting value of 25 (would be 0 otherwise)
N[1] = 25

#Model it starting at point 2 since we already gave a value for point 1
for (t in 2:nsteps){
  r <- rnorm(n = 1, mean = mu, sd = sigma)
  N[t] <-  N[t-1] + r + (1 - N[t - 1] / K)
}

# visualize with a plot

plot(N)
```
 Model 3 (Using a matrix)
 
```{r}
# setup model parameters

K <- 50
nsteps <- 50
mu <- .4
sigma <- 2

# setup model objects

N = matrix(data = NA, nrow = nsteps, ncol = 100)
# assign starting value of 25 (would be 0 otherwise)
N[1, ] = 25

#Model it starting at point 2 since we already gave a value for point 1
for (t in 2:nsteps){
  r <- rnorm(n = 100, mean = mu, sd = sigma)
  N[t,] <-  N[t-1] + r + (1 - N[t - 1,] / K)
}

# visualize with a plot

matplot(N, type = "l")
```
 
Now use tidyverse
```{r}
library(tidyverse)

# create df
my_data <- as_tibble(N) %>% 
  mutate(time = 1:nsteps) %>% 
  gather(run, N, -time)

# visualize it 
ggplot(my_data, aes(x = time, y = N)) +
  geom_line(aes(group = run), size = .1) +
  stat_summaryd(geom = "line", fun = mean, color = "red")
```






